---
title: "P8105 Homework 3"
author: "Ruiyang Li"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

This is my solution to HW3.

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r instacart_df}
data("instacart")
```

Instacart is an online grocery service that provides online shopping from local stores. For example, in New York City, such local partner stores include Whole Foods, Fairway, and The Food Emporium. Normally, orders will be delivered within 2 hours after placement. 

The Instacart data consists of `r nrow(instacart)` rows of online grocery orders and `r ncol(instacart)` variables from `r length(unique(instacart$user_id))` Instacrat customers. In this dataset, each customer has one or more orders, so the data is in the long format. Some of the key variables include information on customers / orders such as user ID, order ID, order day, and order hour, and information on products such as name, aisle, and department. 


* There are `r length(unique(pull(instacart, aisle_id)))` aisles. The most items are ordered from `r instacart %>% count(aisle) %>% arrange(desc(n)) %>% slice(1) %>% select(aisle) %>% as.character()` aisles, which have `r instacart %>% count(aisle) %>% arrange(desc(n)) %>% slice(1) %>% select(n)` orders. 

```{r collapse=TRUE}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


* The following plot shows the aisles with more than 10000 items ordered. 

```{r collapse=TRUE}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

* The next table shows the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. 

```{r collapse=TRUE}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  arrange(desc(n)) %>% 
  slice(1:3) %>% 
  ungroup() %>% 
  knitr::kable()
```

* The table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

```{r collapse=TRUE}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hr = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hr"
  ) %>% 
  knitr::kable()
```

“The Instacart Online Grocery Shopping Dataset 2017”, Accessed from https://www.instacart.com/datasets/grocery-shopping-2017 on June 24, 2017. 


## Problem 2

* Load, tidy, and wrangle the data. 

```{r collapse=TRUE}
accel_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>% 
	pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_", 
    values_to = "activity"
  ) %>%
  mutate(
  	minute = as.numeric(minute), 
    day_cat = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"),
		day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
	relocate(week, day_id, day, day_cat)
```

This tidied accelerometer dataset contains `r nrow(accel_df)` observations and `r ncol(accel_df)` variables. These variables include week, day, minute and activity counts for each minute of a 24-hour day starting at midnight. 

* The following table shows the total activities for each day. There are much fewer activities on the Saturdays of week 4 and 5. 

```{r collapse=TRUE}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(activitiy_total = sum(activity)) %>% 
  knitr::kable()
```

* Here is a single-panel plot that shows the 24-hour activity time courses for each day, colored by day of the week. Based on this graph, it can be seen that Sunday has more activities in the middle of the day; Monday, Wednesday, Friday and Saturday have more activities close to the end of the day; activities are the lowest at the beginning and the end of a day. 

```{r}
accel_df %>% 
	group_by(day, minute) %>% 
  ggplot(aes(x = minute, y = activity, color = day)) + 
	geom_point(alpha = 0.3) + 
	geom_line() + 
	labs(
    title = "24-hour activity time courses",
    x = "Minute of a 24-hour day starting at midnight",
    y = "Activity counts for each minute")
```




